%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%  references.tex
%
%  Bibliography
%
%  $Id: references.tex 1902 2013-01-31 15:11:10Z ohenrich $
%
%  Edinburgh Soft Matter and Statistical Physics Group and
%  Edinburgh Parallel Computing Centre
%
%  (c) 2011-2016 The University of Edinburgh
%  Kevin Stratford (kevin@epcc.ed.ac.uk)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\vfill
\pagebreak

\addcontentsline{toc}{section}{References}

\bibliographystyle{plain}
\begin{thebibliography}{99}


\bibitem{adhikari-desplat2005}
Adhikari, R. J.-C. Desplat, and K. Stratford,
Sliding periodic boundary conditions for lattice Boltzmann and lattice
kinetic equations,
\texttt{arXiv:cond-mat/0503175v1} (2005).

\bibitem{apacheAPR} 
Apache Portable Runtime Project ``APR's Version Numbering''.
See, e.g., \texttt{https://apr.apache.org/versioning.html)}
(accessed November 2015).

\bibitem{adhikari2005}
Adhikari, R., K. Stratford. M.E. Cates, and A.J. Wagner,
Fluctuating Lattice Boltzmann,
\textit{Europhys. Lett.}, \textbf{71}, 473 2005.

\bibitem{ald98}
Aidun, C.K., Y. Lu, and E.-J. Ding,
Direct analysis of particulate suspensions with inertia using the
discrete Boltzmann equation,
\textit{J. Fluid Mech.}, \textbf{373}, 287, 1998.

\bibitem{allen-tildesley}
M.P. Allen and D.J. Tildesley,
\textit{Computer simulation of liquids},
Oxford University Press (1987).

\bibitem{batchelor}
G.K. Batchelor
\textit{An Introduction to Fluid Mechanics},
Cambridge University Press (1967).

\bibitem{beris-edwards}
Beris, A.N., and B.J. Edwards,
\textit{Thermodynamics of flowing systems with internal microstructure},
Oxford University Press (1994).

\bibitem{blake}
J.R. Blake, A spherical envelope approach to ciliary propusion,
\textit{J. Fluid Mech.}, \textbf{46}, 199, 1971.

\bibitem{cahn-hilliard-1958}
J.W. Cahn and J.E. Hilliard,
Free energy of a nonuniform system. I. Interfacial free energy,
\textit{J. Chem. Phys.} \textbf{28} 258--267 (1958).

\bibitem{cates_scaling}
M. E. Cates, J.-C. Desplat, P. Stansell, A.J. Wagner, K. Stratford,
R. Adhikari, and I. Pagonabarraga,
Physical and Computational Scaling Issues in Lattice Boltzmann
Simulations of Binary Fluid Mixtures,
\textit{Phil. Trans. Roy. Soc. A}, \textbf{363}, 1917 (2005). 

\bibitem{chaikin-lubensky}
P.M. Chaikin and T.C. Lubensky,
\textit{Principles of condensed matter physics}, Cambridge University
Press (1995).

\bibitem{chunladd}
B. Chun and A.J.C. Ladd,
Interpolated boundary condition for lattice Boltzmann simulations in
narrow gaps,
\textit{Phys. Rev. E}, \textbf{75}, 066705, 2007.

\bibitem{cj98}
Cichocki, B., and R.B. Jones,
Image representation of a spherical particle near a hard wall,
\textit{Physica A}, \textbf{258}, 273, 1998.

\bibitem{chang}
C.-H. Chang and E.I. Franses,
Adsorption dynamics of surfactants at the air/water interface:
a critical review of mathematical models, data, and mechanisms,
\textit{Colloids and Surfaces A}, \textbf{100} 1, 1995.

\bibitem{degennes-prost2002}
de Gennes, J.-G., and J. Prost,
\textit{The physics of liquid crystals},
Oxford University Press (2002).

\bibitem{desplat2001}
Desplat, J.-C., I. Pagonabarraga, and P. Bladon,
LUDWIG: A parallel lattice-Boltzmann code for complex fluids.
\textit{Comput. Phys. Comms.}, \textbf{134}, 273, 2001.

\bibitem{diamant}
H. Diamant and D. Andelman,
Kinetics of surfactant adsorption at fluid/fluid
interfaces: non-ionic surfactants,
\textit{Europhys. Lett.} \textbf{34}, 575 (1996).

\bibitem{diamant96}
H. Diamant and D. Andelman,
Kinetics of surfactant adsorption at fluid-fluid interfaces,
\textit{J. Phys. Chem.}, \textbf{100} 13732, 1996.

\bibitem{fournier2005}
J.-B. Fournier and P. Galatola,
Modeling planar degenerate wetting and anchoring in nematic liquid
crystals,
\textit{Europhys. Lett.}, \textbf{72} 403--409 (2005).

\bibitem{eastoe}
J. Eastoe and J.S. Dalton,
Dynamic surface tension and adsorption mechanims of surfactants
at the air-water interface,
\textit{Advances in Colloid and Interface Science}, \textbf{85}
13, 2000.

\bibitem{ginzburg}
I. Ginzburg and D. d'Humi\`eres,
Multireflection boundary conditions for lattice Boltzmann models,
\textit{Phys. Rev. E}, \textbf{68}, 066614, 2003.

\bibitem{gray2013}
A. Gray and K. Stratford,
TargetDP reference.

\bibitem{h59}
Hasimoto, H., On the periodic fundamental solutions of the Stokes
equation and their application to viscous flow past a cubic array
of spheres.
\textit{J. Fluid Mech.}, \textbf{5}, 317.

\bibitem{heemels}
Heemels, M.W., M.H.J. Hagen, and C.P. Lowe, Simulating solid colloidal
particles using the lattice-Boltzmann method,
\textit{J. Comp. Phys.}, \textbf{164}, 48, 2000.

\bibitem{ieee-208}
IEEE Standard 208-2005, IEEE Standard for Software Configuration Management
Plans.
See \texttt{http://ieeexplore.ieee.org/xpl/standards.jsp}
(accessed November 2015).

\bibitem{ieee-828-2012}
IEEE Standard 828-2012. IEEE Standard for Configuration
Management in Systems and Software Engineering. 
See \texttt{http://ieeexplore.ieee.org/xpl/standards.jsp}
(accessed November 2015).

\bibitem{jo84}
Jeffrey, D.J., and Y. Onishi,
Calculation of the resistance and mobility functions for the two
unequal rigid spheres in low-Reynolds-number flow,
\textit{J. Fluid Mech.}, \textbf{139}, 261, 1984.

\bibitem{kendon2001}
Kendon, V.M., M.E. Cates, I. Pagonabarraga, J.-C. Desplat, and
P. Bladon,
Inertial effects in three dimensional spinodal decomposition of
a symmetric binary fluid mixture: A lattice Boltzmann study,
\textit{J. Fluid Mech.}, \textbf{440}, 147 (2001).

\bibitem{l94a}
Ladd, A.J.C., Numerical simulations of particulate suspensions
via a discretised Boltzmann equation. Part 1. Theoretical foundation,
\textit{J. Fluid. Mech.}, \textbf{271}, 285, 1994.

\bibitem{l94b}
Ladd, A.J.C., Numerical simulations of particulate suspensions
via a discretised Boltzmann equation. Part 2. Numerical results,
\textit{J. Fluid. Mech.}, \textbf{271}, 311, 1994.

\bibitem{l96a}
Ladd, A.J.C., Sedimentation of homogenous suspensions of non-Brownian
spheres,
\textit{Phys. Fluids}, \textbf{9}, 491. 1996.

\bibitem{l96b}
Ladd, A.J.C., Hydrodynamic screening in sedimentating suspensions
of non-Brownian spheres,
\textit{Phys. Rev. Lett.}, \textbf{76}, 1392, 1996.

\bibitem{lv01}
Ladd, A.J.C., and R. Verberg,
Lattice-Boltzmann simulations of particle-fluid suspensions,
\textit{J. Stat. Phys.}, \textbf{104}, 1191, 2001.

\bibitem{lees-edwards1972}
A.W. Lees and S.F. Edwards,
The computer study of transport processes under extreme conditions,
\textit{J. Phys C} \textbf{5} 1921--1929 (1972).

\bibitem{lipanmiller}
H. Li, C. Pan, and C.T. Miller,
Pore-scale investigation of viscous coupling effects for two-phase
flow in porous media,
\textit{Phys. Rev. E}, \textbf{72}, 026705, 2005.

\bibitem{lighthill}
M.J. Lighthill,
On the squirming motion of nearly spherical deformable bodies through
liquid at very small Reynolds numbers,
\textit{Comm. Pure Appl. Math.}, \textbf{5}, 109, 1952.

\bibitem{isaac}
I. Llopis Fust\'e,
\textit{Hydrodynamic cooperativity in micro-swimmer suspensions},
Ph.D. Thesis, University of Barcelona, 2008.

\bibitem{mpi-standard}
Message Passing Interface Forum. MPI: A Message Passing Interface Standard
Version 1.3 (2008).

\bibitem{nguyen-ladd2002}
Nguyen, N.-Q., and A.J.C. Ladd, Lubrication corrections for
lattice-Boltzmann simulations of particle suspensions,
\textit{Phys. Rev. E}, \textbf{66}, 046708, 2002.

\bibitem{papanastasiou}
T. paapnastasiou, G. Georgiou, and A. Alexandrou,
\textit{Viscous Fluid Flow},
CRC Press, Boca Raton, Florida, 2000.

\bibitem{paraview}
Paraview. See \texttt{http://www.paraview.org/}. Accessed 2011.

\bibitem{rapaport1995}
Rapaport, D.C., \textit{The Art of Molecular Dynamics Simulation},
Cambridge University Press, 1995.


\bibitem{j-stat-phys-2005}
K. Stratford, R. Adhikari, I. Pagonabarraga, and J.-C. Desplat,
Lattice Boltzmann for binary fluids with suspended colloids,
\textit{J. Stat. Phys.} \textbf{121} 163--178 (2005).

\bibitem{succi}
S. Succi, \textit{The lattice Boltzmann equation and beyond},
Oxford University Press, Oxford, 2001.

\bibitem{swift1996}
M.R. Swift, E. Orlandini, W.R. Osborn, and J.M. Yeomans,
Lattice Boltzmann simulation of liquid-gas and binary fluid systems,
\textit{Phys. Rev. E}, \textbf{54} 5041 (1996).

\bibitem{edo1}
M. Venuroli and E.S. Boek,
Two-dimensional lattice-Boltzmann simulations of single phase
flow in a pseudo two-dimensional micromodel,
\textit{Physica A}, \textbf{362}, 23, 2006.

\bibitem{vandergraaf}
R.G.M. van der Sman and S. van der Graaf,
Diffuse interface model of surfactant adsorption onto flat and
droplet interfaces,
\textit{Rheol. Acta} \textbf{46} 3 (2006).

\bibitem{theissengompper}
O. Theissen and G. Gompper,
Lattice Boltzmann study of spontaneous emulsification,
\textit{Eur. Phys. J. B}, \textbf{11} 91 (1999).

\bibitem{wagner-pagonabarraga2002}
A.J. Wagner and I. Pagonabarraga,
Lees-Edwards boundary conditions of lattice Boltzmann,
\textit{J. Stat. Phys.} \textbf{107} 521--537 (2002).

\bibitem{wardtordai}
A.F.H. Ward and L. Tordai,
\textit{J. Chem. Phys.} \textbf{14} 453, 1946.

\bibitem{skarabot}
M. Skarabot, M. Ravnik, S. Zumer, U. Tkalec, I. Poberaj, D. Babic, N. Osterman and I. Musevic,
\textit{Phys. Rev. E} \textbf{76}, 051406 (2007).

\bibitem{wright-mermin}
Wright, D.C> and N.D. Mermin,
\textit{Rev. Mod. Phys.} \textbf{61}, 385--432 (1989).

\bibitem{Lyklema} J. Lyklema {\em Fundamentals of Interface and Colloid Science} Academic Press     (1995).
\bibitem{Mafe} S. Maf\'e, J.A. Manzanares, J. Pellicer, {\textit J. Electroanal. Chem.} {\textbf 241}, 5    7-77 (1988).
\bibitem{Capuani} F. Capuani, I. Pagonabarraga, D. Frenkel, {\textit J. Chem. Phys.} {\textbf 121}, 973-    986 (2004).
\bibitem{Rotenberg} B. Rotenberg, I. Pagonabarraga, D. Frenkel, {\textit Farad. Discuss.} {\textbf 144},     223-243 (2010).
\bibitem{Landau-ED} L.D. Landau, E.M. Lifshitz, {\textit Electrodynamics of Continuous Media}, \S 15    , 2nd ed., Pergamon Press, Oxford, UK (1984).
\bibitem{Melcher} J.R. Melcher, {\textit Continuum Electromechanics}, \S 3.10, MIT Press, Cambridge,     MA, USA (1981).\\
downloadable from:\\
\url{http://ocw.mit.edu/ans7870/resources/melcher/resized/cem_811.pdf}
\bibitem{Landau-EL} L.D. Landau, E.M. Lifshitz, {\textit Theory of Elasticity}, \S 3 \& \S 16, 3rd e    d., Butterworth-Heinemann, Oxford, UK (1986).



\end{thebibliography}


